### 2018-03-21 tex/paper.tex
Add example and note about Money to paper

\noindent
\sjtnote{Preliminary overview; need to revisit.}
The paper is structured like this: we begin with an overview of the model, including the assumptions made in designing it, the types of the principal functions and a description of the DSL as an algebraic type, constructor by constructor. The section aims to give a high level overview, and some of the concrete details are to be found in the repo. We follow this with discussions of the kinds of formal analysis which might be made for the DSL, as well as discussing a  set of questions raised by the process of modelling. We then look at the example of BLAH, as an illustration, and then look at how Marlowe relates to other work in the area. We conclude by pointing out

@@

\begin{itemize}
\item It is assumed that cryptographic ``services'' or operations are provided by a layer external to the system, and so they need not be modelled explicitly here.
\item Making a commitment is not something that a contract can perform; rather, it can require that a commitment is made, but that then has to be established externally: hence the input of (a set of) commitments at each step. The same observation also applies to redeeming a commitment of cash.
\item Observables are observed at a particular time and in a particular context. We assume that the recording and identification of each (instance of an) observation is provided by the infrastructure of the system. We provide a ``little language'' for describing observations: \haskellinline{Observation}.
\item The fact of making commitments also needs to be recorded.
\item It is important to be aware that there may be different instances of contracts executed, and for each instance there will typically be different instances of observables (and commitments) observed and recorded.
\item The model manages the release of funds back to the committer when a cash commitment expires (see discussion of the \haskellinline{fullStep} function below).
\item We assume that time is ``coarse grained'' and measured simply by block number, so that timeouts etc.\ are delimited using nothing finer than block number. 
\end{itemize}

@@

\begin{haskellcode}
data Contract =
   Null |
   CommitCash IdentCC Person Cash Timeout Timeout Contract Contract |  
   RedeemCC IdentCC Contract |
   Pay IdentPay Person Person Cash Timeout Contract |  
   Both Contract Contract |
   Choice Observation Contract Contract |
   When Observation Timeout Contract Contract   
   \end{haskellcode}
%Originally on "Choice": \footnote{Doesn't the step function need to record what observations are made, so they can be archived in the blockchain? I don't see how your design will support doing so.}\footnote{Yes, we assume that it is part of the infrastructure of this language to record the values of *observables* that are used. However, we have chosen no to make it part of the language itself.}

@@

\item \haskellinline{CommitCash idCC p n t1 t2 k1 k2}  For this contract to make progress,\begin{itemize}
\item either before the timeout \haskellinline{t1} the user \haskellinline{p} makes a cash commitment of value \haskellinline{n} and timeout \haskellinline{t2}  with the identifier \haskellinline{idCC} , 
\item
or the timeout \haskellinline{t1} is exceed.
\end{itemize} 
In the first case, a \haskellinline{SuccessfulCommit} action is generated and the contract continues as 
\haskellinline{k1}; in the second case no action is generated and the contract continues as 
\haskellinline{k2}. While neither case holds the contract is quiescent, waiting for the cash to be committed.

@@

\item \haskellinline{Pay p1 p2 int t k} makes it possible, assuming that sufficient funds are available, for \haskellinline{p1} to claim  a payment of \haskellinline{int} from \haskellinline{p2} before the timeout \haskellinline{t}. The contract continues as \haskellinline{k}. By `available' we mean that sufficient commitments have been made and  not yet expired to cover the payment; in this case the payment  uses the currency allocated by the cash commitments made by \haskellinline{p1} that expire the earliest.

@@

\section{Rock-paper-scissors explanation (not re-formatted)}

In this section, we explain how the DSL can be used for designing a contract that enforces a round of rock-paper-scissors where the winner will receive from the loser 10 units of currency and not revealing your hand also has a penalty of 10 units.

We will start top to bottom. First we define identities for the players and the commitments.
The player 1 will be identified by integer 1:
player1 = 1
the cash commitment of player 1 will be identified by:
p1 = IdentCC 1
and his value commitment will be identified by:
v1 = IdentCV 1
In the same way, player 2 will be identified by integer 2:
player2 = 2
the cash commitment player 2 will be identified by:
p2 = IdentCC 2
And his value commitment will be identified by:
v2 = IdentCV 2

And because choices must be represented as Int we encode them as follows:

data RPS = Rock | Paper | Scissors

all\_rps = [Rock, Paper, Scissors]

encoding\_rps Rock = 1      -- rock is encoded as 1
encoding\_rps Paper = 2     -- paper is encoded as 2
encoding\_rps Scissors = 3  -- scissors is encoded as 3

all\_rps\_encoded = map (encoding\_rps) all\_rps

We first ask the players to commit money. We will need 20 units from each player since we want 10 units to enforce that the hand is revealed, and 10 units for the prize. Theoretically we could reuse just 10 units, but in practice recommitting payments may limit the applicability of the DSL to certain systems like Bitcoin, so it does not support it at the moment.

The commitments are chained as follows:
* Cash commitment p1 is done by player 1, consists of 20 units and expires on block 42 same for p2 by player 2.

CommitCash p1 player1 20 42
  (CommitCash p2 player2 20 42

* Value commitments are done for two choices of one of the three values that encode rock, paper, or scissors.

     (CommitValue v1 (all\_rps\_encoded) player1
        (CommitValue v2 (all\_rps\_encoded) player2

* The see-saw mechanism ensures that both players reveal or none does. This is fundamentally achieved by forcing a player to reveal the other when he reveals. But we need to do this steadily to give players enough time to do the actions required by the rest of the contract, otherwise one player may push the other one to timeout.

    (RevealCV v1
            (Choice (BelowTimeout 10)
                    (Pay player2 player1 10
                       (RevealCV v2
                              (Choice (BelowTimeout 20)
                                  (Pay 1 2 10
                                       reward\_winner)
                                  Null)))
                    Null)))))

First RevealCV waits for the owner of value commitment v1 (player 1) to reveal his choice. If the reveal happens before block number 10 (after the start of the contract instance), player 2 pays 10 currency units to player 1. Otherwise we consider the contract timed out and make the contract Null, the existing cash commitments will be automatically refunded on block 42.
The same is done for player 2. This time he is incentivised to reveal his hand before block number 20, since that is the only way he will recover the 10 currency unit. He might already know that he has lost and decide not to reveal, that will have the same effect than revealing. If desired we could increase the see-saw penalty to make not revealing worse than losing.

If the contract remaining becomes reward\_winner, we already have enough information to know who won and we will have exactly 10 currency units committed by each player.

So we define reward\_winner as follows. We just use Choice twice, one to check if Player 1 won, one to check if Player 2 won, and the default option is just refunding everything:

-- If player 1 won
Choice (OrObs (AndObs (CVRevealedAs v1 (encoding\_rps Paper))
                      (CVRevealedAs v2 (encoding\_rps Rock)))
              (OrObs (AndObs (CVRevealedAs v1 (encoding\_rps Scissors))
                             (CVRevealedAs v2 (encoding\_rps Paper)))
                     (AndObs (CVRevealedAs v1 (encoding\_rps Rock))
                             (CVRevealedAs v2 (encoding\_rps Scissors)))))
-- player2 pays 10 currency units to player 1
       (Pay player2 player1 10
-- and everything else is redeemed
          (Both (RedeemCC p1 Null)
                (RedeemCC p2 Null)))
-- if player 2 won
       (Choice (OrObs (AndObs (CVRevealedAs v2 (encoding\_rps Paper))
                              (CVRevealedAs v1 (encoding\_rps Rock)))
                      (OrObs (AndObs (CVRevealedAs v2 (encoding\_rps Scissors))
                                     (CVRevealedAs v1 (encoding\_rps Paper)))
                             (AndObs (CVRevealedAs v2 (encoding\_rps Rock))
                                     (CVRevealedAs v1 (encoding\_rps Scissors)))))
-- player1 pays 10 currency units to player 2
               (Pay player1 player2 10
-- and everything else is redeemd
                  (Both (RedeemCC p1 Null)
                        (RedeemCC p2 Null)))
-- otherwise everything else is redeemed
               (Both (RedeemCC p1 Null)
                     (RedeemCC p2 Null)))

It is also worth noting that, because the possible choices are specified, RedeemCV requires that the choice revealed is valid, otherwise it behaves as nothing was revealed. Thus, we know that if no player won it must be a draw.

>>>>

\noindent
\sjtnote{Preliminary overview; need to revisit.}
The paper is structured like this: we begin with an overview of the model, including the assumptions made in designing 
it, the types of the principal functions and a description of the DSL as an algebraic type, constructor by constructor. 
The section aims to give a high level overview, and some of the concrete details are to be found in the repo. We follow 
this with discussions of the kinds of formal analysis which might be made for the DSL, as well as discussing a  set of 
questions raised by the process of modelling. We then look at the example of ``incentivised saving'', as an 
illustration, and then look at how Marlowe relates to other work in the area. We conclude by pointing out

@@

\begin{itemize}
\item It is assumed that cryptographic ``services'' or operations are provided by a layer external to the system, and so they need not be modelled explicitly here.
\item Making a commitment is not something that a contract can perform; rather, it can require that a commitment is made, but that then has to be established externally: hence the input of (a set of) commitments at each step. The same observation also applies to redeeming a commitment of cash.
\item Observables are observed at a particular time and in a particular context. We assume that the recording and 
identification of each (instance of an) observation is provided by the infrastructure of the system. We provide a 
``little language'' for describing observations: \haskellinline{Observation}. Similar properties are true for
\haskellinline{Money} amounts: their values are calculated depending on the inputs and state of execution of the 
contract, they can be deterministically established given the context, and they are described through a small 
``expression language''.
\item The fact of making commitments also needs to be recorded.
\item It is important to be aware that there may be different instances of contracts executed, and for each instance there will typically be different instances of observables (and commitments) observed and recorded.
\item The model manages the release of funds back to the committer when a cash commitment expires (see discussion of the \haskellinline{fullStep} function below).
\item We assume that time is ``coarse grained'' and measured simply by block number, so that timeouts etc.\ are delimited using nothing finer than block number. 
\end{itemize}

@@

\begin{haskellcode}
data Contract =
   Null |
   CommitCash IdentCC Person Money Timeout Timeout Contract Contract |  
   RedeemCC IdentCC Contract |
   Pay IdentPay Person Person Money Timeout Contract |  
   Both Contract Contract |
   Choice Observation Contract Contract |
   When Observation Timeout Contract Contract   
   \end{haskellcode}
%Originally on "Choice": \footnote{Doesn't the step function need to record what observations are made, so they can be archived in the blockchain? I don't see how your design will support doing so.}\footnote{Yes, we assume that it is part of the infrastructure of this language to record the values of *observables* that are used. However, we have chosen no to make it part of the language itself.}

@@

\item \haskellinline{CommitCash idCC p v t1 t2 k1 k2}  For this contract to make progress,\begin{itemize}
\item either before the timeout \haskellinline{t1} the user \haskellinline{p} makes a cash commitment of value 
\haskellinline{v} and timeout \haskellinline{t2}  with the identifier \haskellinline{idCC}, 
\item
or the timeout \haskellinline{t1} is exceed.
\end{itemize} 
In the first case, a \haskellinline{SuccessfulCommit} action is generated and the contract continues as 
\haskellinline{k1}; in the second case no action is generated and the contract continues as 
\haskellinline{k2}. While neither case holds the contract is quiescent, waiting for the cash to be committed.

@@

\item \haskellinline{Pay p1 p2 v t k} makes it possible, assuming that sufficient funds are available, for 
\haskellinline{p1} to claim  a payment of \haskellinline{v} from \haskellinline{p2} before the timeout 
\haskellinline{t}. The contract continues as \haskellinline{k}. By `available' we mean that sufficient commitments have 
been made and  not yet expired to cover the payment; in this case the payment  uses the currency allocated by the cash 
commitments made by \haskellinline{p1} that expire the earliest.

@@

\section{Incentivised saving}

In this section, we explain how the DSL can be used for designing a contract that incentivises saving. We use the 
following example: ``participant 2'' offers 20 ADA to ``participant 1'' in exchange for saving 100 ADA for 100 blocks. 
``Participant 1'' will be able to withdraw the savings after block 20, but if that option is chosen, then ``participant 
2'' will not pay the extra 20 ADA to ``participant 1''.

We will start from outer to inner primitives. The first thing is to allow ``participant 1'' to commit the 100 ADA. The 
primitive says that with identity \haskellinline{IdentCC 1} (this is the identity of the commitment), ``participant 
\haskellinline{1}'' can commit \haskellinline{100} ADA before block \haskellinline{10}, and the commitment must 
expire on block \haskellinline{200}. But if the money is not committed before block \haskellinline{10}, then the 
contract will continue as \haskellinline{Null}.

\begin{haskellcode}
CommitCash (IdentCC 1) 1 (ConstMoney 100) 10 200
           step2
           Null
\end{haskellcode}

If the commitment is done correctly and on time, we continue as in \haskellinline{step2}. \haskellinline{step2} is 
another commitment, this time by ``participant 2''. The primitive says that with identity \haskellinline{IdentCC 2},
``participant \haskellinline{2}'' can commit \haskellinline{20} ADA before block \haskellinline{20}, and the commitment
must expire on block \haskellinline{200}. But if the money is not committed before block \haskellinline{20}, then the
contract will continue as \haskellinline{RedeemCC (IdentCC 1) Null}, which in turn gives ``participant 1'' the 
opportunity to withdraw all the money committed immediately.

\begin{haskellcode}
CommitCash (IdentCC 2) 2 (ConstMoney 20) 20 200
           step3
           (RedeemCC (IdentCC 1) Null)
\end{haskellcode}

If the commitment is done correctly and on time, we continue as in \haskellinline{step3}. \haskellinline{step3}
waits until choice with identity \haskellinline{IdentChoice 1} is made by ``participant \haskellinline{1}'' or
until block \haskellinline{100} (whatever happens first). If the choice is made before block \haskellinline{100}, 
it continues as \haskellinline{stepRedeem}, if block \haskellinline{100} comes first, then it continues as 
\haskellinline{step4}.

\begin{haskellcode}
When (PersonChoseSomething (IdentChoice 1) 1) 100
     stepRedeem
     step4
\end{haskellcode}

\haskellinline{step4} pays the \haskellinline{20} ADA committed by ``participant \haskellinline{2}'' in the commit with 
identity \haskellinline{IdentPay 1} to ``participant \haskellinline{1}'', if the payment is claimed before block 
\haskellinline{200}. Once the payment is claimed or after block 200, it continues as \haskellinline{stepRedeem}.

\begin{haskellcode}
Pay (IdentPay 1) 2 1 (ConstMoney 20) 200 stepRedeem
\end{haskellcode}

Finally, \haskellinline{stepRedeem} allows both participants to redeem all the remaining committed funds.

\begin{haskellcode}
Both (RedeemCC (IdentCC 1) Null)
     (RedeemCC (IdentCC 2) Null)
\end{haskellcode}

The whole contract remains as follows:
\begin{haskellcode}
CommitCash (IdentCC 1) 1
           (ConstMoney 100)
           10 200
           (CommitCash (IdentCC 2) 2
                       (ConstMoney 20)
                       20 200
                       (When (PersonChoseSomething (IdentChoice 1) 1)
                             100
                             (Both (RedeemCC (IdentCC 1) Null)
                                   (RedeemCC (IdentCC 2) Null))
                             (Pay (IdentPay 1) 2 1
                                  (ConstMoney 20)
                                  200
                                  (Both (RedeemCC (IdentCC 1) Null)
                                        (RedeemCC (IdentCC 2) Null))))
                       (RedeemCC (IdentCC 1) Null))
           Null
\end{haskellcode}
